---
title: "IST687_Week5_DataMunging"
author: "Brandon Croarkin"
date: "May 11, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Munging

Can connect to a large array of datasets already included in R

```{r}
#can see all the datasets that are included in R
data()
```

```{r}
#going to briefly explore a dataset included in R
head(BOD)
```

```{r}
summary(BOD)
```

```{r}
str(BOD)
```

```{r}
help(BOD)
```

####Working with Excel data

```{r}
#opening the file
df <- read_excel("nst-est2011-01.xls")
```

```{r}
#opening the file
df2 <- read_excel("nst-est2011-01.xls")
```

```{r}
#examining the file
str(df)
```

Clearly we need to clean this dataset
```{r}
#remove the header rows
df <- df[-1:-8,]
head(df)
```

```{r}
#keep only the first five problems
df <- df[,1:5]
tail(df)
```

```{r}
#remove the last five rows
df <- df[-52:-62,]
tail(df)
```

```{r}
#give our column header a better name. 
colnames(df)[1] <- "stateName"
```

```{r}
#clean the format of the states in column 1
df$stateName <- str_replace(df$stateName,"\\.","")
head(df)
```

```{r}
#let's make a function to get rid of commas, spaces, and convert to number
Numberize <- function(inputVector) {
  #get rid of commas
  inputVector <- gsub(",","",inputVector)
  #get rid of spaces
  inputVector <- gsub(" ","",inputVector)
  #convert to number
  inputVector <- as.numeric(inputVector)
}
```

```{r}
#change names and clean up the columns with numbers
df$april10census <- Numberize(df$X__1)
df$april10base <- Numberize(df$X__2)
df$july10pop <- Numberize(df$X__3)
df$july11pop <- Numberize(df$X__4)
head(df)
```

```{r}
#need to delete the columns we duplicated/cleaned-up
df <- df[,-2:-5]
head(df)
```
File is cleaned up and ready for analysis!

####Working with Databases

```{r}
#establish R connection to SQL
conSQL <- odbcConnect("ist659_project")
```

```{r}
#assign SQL table list
tblsSQL <- sqlTables(conSQL)
tblsSQL
```

```{r}
dataSQL <- sqlQuery(conSQL,paste("select * from Employee"))
dataSQL
```

```{r}
#need to set sqldf driver to another option since it was previously looking at MySQL
options(sqldf.driver = 'SQLite')
results <- sqldf('select mtcars.mpg from mtcars')
results
```

```{r}
mean(results$mpg)
```

```{r}
results <- sqldf('select mtcars.mpg from mtcars where cyl=4')
results
```

####JSON Data

```{r}
MakeGeoURL <- function(address) {
  root <- "http://maps.google.com/maps/api/geocode/"
  url <- paste(root, "json?address=",address,"&sensor=false",sep = "")
  return(URLencode(url))
}
```

```{r}
MakeGeoURL("1600 Pennsylvania Avenue,Washington, DC")
```

```{r}
Addr2latlng <- function(address){
  url <- MakeGeoURL(address)
  apiResult <- getURL(url)
  geoStruct <- fromJSON(apiResult,simplify=FALSE)
  lat <- NA
  lng <- NA
  try(lat <- geoStruct$results[[1]]$geometry$location$lat)
  try(lng <- geoStruct$results[[1]]$geometry$location$lng)
  return(c(lat,lng))
}
```

```{r}
#test function with the White House
testData <- Addr2latlng("1600 Pennsylvania Avenue,Washington, DC")
str(testData)
```

```{r}
#test function with my address
testData <- Addr2latlng("21284 Arcadia Court, Ashburn, VA")
str(testData)
```

So cool! Let's try accesssing different JSON data.

```{r}
bikeURL <- "https://feeds.citibikenyc.com/stations/stations.json"
apiResult <- getURL(bikeURL)
```
If you run the head on apiResult you get a crazy long output!

```{r}
results <- fromJSON(apiResult)
str(results)
```

```{r}
results[1]
results[2]
```

Can see that all the actual data is stored in results[2].

```{r}
allBikeData <- results$stationBeanList
str(allBikeData)
```

```{r}
head(allBikeData)
```

```{r}
allBikeData[1,]
```

```{r}
nrow(allBikeData)
```

```{r}
mean(allBikeData$availableDocks)
```

The function below is showing that there are 788 docks that have more than one dock available. 
```{r}
tapply(allBikeData$availableDocks,allBikeData$availableDocks>0,length)
```

```{r}
bikedf <- data.frame(matrix(unlist(allBikeData),nrow=810,byrow=T),stringsAsFactors = FALSE)
head(bikedf)
```


####Sapply and Tapply

```{r}
sapply(mtcars, mean)
```

```{r}
tapply(mtcars$mpg, mtcars$cyl, mean)
```

```{r}
sqldf('SELECT AVG(mtcars.mpg) FROM mtcars GROUP BY mtcars.cyl')
```

```{r}
tapply(mtcars$mpg, mtcars$cyl==4,mean)
```

```{r}
sqldf('SELECT AVG(mtcars.mpg) FROM mtcars WHERE mtcars.cyl = 4')
```

```{r}
goodMPGcar <- function(mpg){
  if(mpg > 24){
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

```{r}
sapply(mtcars$mpg,goodMPGcar)
```

```{r}
sum(sapply(mtcars$mpg,goodMPGcar))
```

```{r}
goodCars <- sapply(mtcars$mpg,goodMPGcar)
mtcars[goodCars,]
```


















