---
title: "GGPLOT"
author: "Brandon Croarkin"
date: "May 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GGPLOT 2 Practice

```{r}
library(ggplot2)
```

```{r}
df <- mtcars
```

```{r}
head(df)
```

An example of the basic graphics we can make with R
```{r echo=FALSE}
hist(df$mpg,breaks=4)
```

Now let's try to re-create this with ggplot
```{r echo=FALSE}
g <- ggplot(df,aes(x=mpg)) + geom_histogram(bins=5)
g
```

It is similar, but it actually a little worse than the one before. Let's try to make it better.

```{r echo=FALSE}
g <- ggplot(df,aes(x=mpg)) + geom_histogram(bins=5, color="black",fill="white")
g 
```

```{r echo=FALSE}
g + ggtitle("MPG Buckets")
```


```{r echo=FALSE}
ggplot(df, aes(x=mpg)) + geom_histogram(binwidth = 5, color="black",fill="blue")
```

```{r echo=FALSE}
ggplot(df, aes(x=factor(0),mpg)) + geom_boxplot()
#adding factor(0) lumps all the mpg together
```

```{r echo=FALSE}
ggplot(df, aes(group=cyl, x=cyl,mpg)) + geom_boxplot()
#adding factor(0) lumps all the mpg together
```


```{r echo=FALSE}
ggplot(df, aes(group=cyl, x=cyl,mpg)) + geom_boxplot() + coord_flip()
```

```{r echo=FALSE}
ggplot(df, aes(x=cyl)) + geom_bar()
```

```{r echo=FALSE}
ggplot(df, aes(x=cyl)) + geom_bar()
```

```{r echo=FALSE}
car.names <- rownames(df)
```

```{r echo=FALSE}
g <- ggplot(df, aes(x=car.names, y=wt)) + geom_bar(stat="identity")
g
```

Need to clean up the x-axis to make it readable.

```{r echo=FALSE}
g <- g + theme(axis.text.x = element_text(angle=90, hjust=1))
g
```

Looks better! Let's just clean it up a little more. 

```{r echo=FALSE}
g + ggtitle("Car Weights") + theme(plot.title = element_text(hjust=.5))
```

Let's make a stacked bar chart. 

```{r echo=FALSE}
#the fill is a factor of gears
ggplot(df, aes(x=cyl, fill=factor(gear))) + geom_bar()
```

We can also unstack this bar chart

```{r echo=FALSE}
#the fill is a factor of gears
ggplot(df, aes(x=cyl, fill=factor(gear))) + geom_bar(position="dodge")
```

Let's play with some scatters.

```{r echo=FALSE}
ggplot(df, aes(x=mpg,y=hp)) + geom_point()
```

Let's build this up a little.
```{r echo=FALSE}
ggplot(df, aes(x=mpg,y=hp)) + geom_point(aes(size=wt,color=cyl))
```

```{r echo=FALSE}
ggplot(df, aes(x=factor(0),mpg)) + geom_boxplot()
```

```{r echo=FALSE}
ggplot(df, aes(x=factor(0),mpg)) + geom_boxplot()
```

















