---
title: "Week3Practice - Functions"
author: "Brandon Croarkin"
date: "April 26, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Writing Functions

This will run through how to create functions in R. Our goal is to re-create the Mode function to find the most frequently occuring number in an array

```{r}
#First we will create some data to work with
tinyData <- c(1,2,1,2,3,3,3,4,5,4,5)
tinyData

#Then we can create the framework for the function
myMode <- function(myVector) {
  uniqueValues <- unique(myVector)
  return(uniqueValues)
}

#Can test this function out on the tinyData
myMode(tinyData)
```

Now we need to edit this formula to work as intended. 

```{r}
#tabulate gives the count for each unique variable
tabulate(tinyData)

#apply to our function
myMode <- function(myVector) {
  uniqueValues <- unique(myVector)
  uniqueCounts <- tabulate(myVector)
  return(uniqueVector)
}

dataDisplay <- data.frame(unique(tinyData), tabulate(tinyData))
dataDisplay
```

We can see that we now can create a data frame with the unique number and the count of each number. We now need to pick the unique number that has the highest count. 

```{r}
myMode <- function(myVector) {
  uniqueValues <- unique(myVector)
  uniqueCounts <- tabulate(myVector)
  return(uniqueValues[which.max(uniqueCounts)])
}

tinyData
myMode(tinyData)

```

Now we want to test the function some more to see if it has any flaws/errors.

```{r}
tinyData <- c(tinyData, 9,9,9,9,9,9,9)
myMode(tinyData)
```

Can see we are getting an NA. Why?
```{r}
str(tinyData)
summary(tinyData)
tabulate(tinyData)
max(tabulate(tinyData))
unique(tinyData)
max(c(5,2,3,2,0,0,0,7))
```

The problem is that the two arrays (uniqueValues and uniqueCounts) were not of the same length. This is because it tabulate gave counts the missing numbers (6,7,8)
```{r}
myMode <- function(myVector) {
  uniqueValues <- unique(myVector)
  uniqueCounts <- tabulate(match(myVector,uniqueValues))
  return(uniqueValues[which.max(uniqueCounts)])
}

myMode(tinyData)
```

SUCCESS!! The match function only tabulated where there was a match between the vectors of tinyData and the uniqueValues in tinyData.

Note: in R the function to do the mode of a data set is mfv
```{r}
library(modeest)
mfv(tinyData)
```