---
title: "Croarkin_Week5HW"
author: "Brandon Croarkin"
date: "May 12, 2018"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## IST687
### JSON & tapply HW: Accident Analysis

####Step 1: Load the data

```{r}
library(jsonlite)
library(sqldf)
library(RCurl)
```

```{r}
URL <- "http://data.maryland.gov/api/views/pdvh-tf2u/rows.json?accessType=DOWNLOAD"
apiResult <- getURL(URL)
```

```{r}
json_file <- fromJSON(apiResult)
```

Looks like all the data is in the second element.

```{r}
data <- json_file[[2]]
```

```{r}
numRows <- nrow(data)
numRows
```

```{r}
df <- data.frame(data)
```

####Step 2: Clean the data

```{r}
#need to remove the first 8 columns
dfClean <- df[,-1:-8]
head(dfClean)
```

```{r}
namesOfColumns <-
c("CASE_NUMBER","BARRACK","ACC_DATE","ACC_TIME","ACC_TIME_CODE","DAY_OF_WEEK","ROAD","INTERSECT_ROAD","DIST_FROM_INTERSECT","DIST_DIRECTION","CITY_NAME","COUNTY_CODE","COUNTY_NAME","VEHICLE_COUNT","PROP_DEST","INJURY","COLLISION_WITH_1","COLLISION_WITH_2")
```

```{r}
names(dfClean) <- namesOfColumns
head(dfClean)
```

```{r}
#remove spaces from DAY_OF_WEEK column
dfClean$DAY_OF_WEEK <- gsub(" ", "",dfClean$DAY_OF_WEEK)
```

####Understand the data using SQL (via SQLDF)

```{r}
#testing sqldf
sqldf('SELECT * FROM dfClean')
```

Question 1: How many accidents happen on SUNDAY?
```{r}
sqldf('SELECT COUNT(*) FROM dfClean WHERE DAY_OF_WEEK="SUNDAY"')
```
2373 accidents happen on Sunday. 

Question 2: How many accidents had injuries?

```{r}
#explore whether there are NA's in the injury column
unique(dfClean$INJURY)
```
There are NA values!

```{r}
#let's make a new df
dfClean2 <- dfClean[-which(is.na(dfClean$INJURY)),]
unique(dfClean2$INJURY)
```

```{r}
sqldf('SELECT COUNT(*) FROM dfClean2 WHERE INJURY="YES"')
```
6433 accidents had injuries.

Question 3: List the injuries by day
```{r}
sqldf('SELECT DAY_OF_WEEK, COUNT(*) FROM dfClean2 WHERE INJURY="YES" GROUP BY DAY_OF_WEEK')
```
It looks like most injuries (1043) happen on Friday.

####Understand the data using tapply

Question 1: How many accidents happen on SUNDAY?
```{r}
tapply(dfClean2$CASE_NUMBER,dfClean2$DAY_OF_WEEK=="SUNDAY",length)
```
2373 accidents happen on Sunday. This matches what we have above with sqldf.

Question 2: How many accidents had injuries?
```{r}
tapply(dfClean2$CASE_NUMBER,dfClean2$INJURY,length)
```
6433 accidents had injuries.This matches what we have above with sqldf.

Question 3: List the injuries by day
```{r}
tapply(dfClean2$INJURY=="YES",dfClean2$DAY_OF_WEEK,sum)
```
This matches what we have above with sqldf!





