---
title: "Croarkin_Week8HW"
author: "Brandon Croarkin"
date: "June 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gdata)
library(ggplot2)
library(relaimpo)
```

## IST 687 - Making Predictions

The	textbook's	chapter	on	linear	models ("Line	Up,	Please")	introduces	linear	predictive	modeling	using	the	workhorse	tool	known	as	multiple	regression.	The	term	"multiple	regression"	has	an	odd	history,	dating	back	to	an	early	scientific	observation	of	a	phenomenon	called	"regression	to	the	mean."	These	days,	multiple	regression	is	just	an	interesting	name	for	using	a	simple	linear	modeling	technique	to	measuring	the	connection	between	one	or	more	predictor	variables	and	an	outcome	variable. In	this	exercise,	we	are	going	to	use	an	open	data	set	to	explore	antelope	population.

This	is	the	first	exercise	of	the	semester	where	there	is	no	sample	R	code	to	help	you	along.	Because	you	have	had	so	much	practice	with	R	by	now,	you	can	create	and/or	find	all	of	the	
code	you	need	to	accomplish	these	steps:

1. Read	in	data	from	the	following	URL:	
http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/excel/mlr01.xls

This	URL	will	enable	you	to	download	the	dataset	into	excel.	
The	more	general	web	site	can	be	found	at:
http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/frames/frame.html

If	you	view	this	in	a	spreadsheet,	you	will	find	that	four	columns	of	a	small	dataset. The	first	column	shows	the	number	of	fawn	in	a	given	spring	(fawn	are	baby	Antelope).	The	second	column	shows	the	population	of	adult	antelope,	the	third	shows	the	annual	precipitation	that	year,	and	finally,	the	last	column	shows	how	bad	the	winter	was	during	that	year.

```{r}
URL <- "http://college.cengage.com/mathematics/brase/understandable_statistics/7e/stud
ents/datasets/mlr/excel/mlr01.xls"
```

2. You	have	the	option	of	saving	the	file	save	this	file	to	your	computer	and	read	it	into	R,	or reading the	data	directly	from	the	web	into	a	data	frame.
```{r}
df <- read.xls(URL)
head(df)
```

3. You	should	inspect	the	data	using	the	str()	command	to	make	sure	that	all	of	the	cases	have	been	read	in	(n=8	years	of observations)	and	that	there	are	four variables.	
```{r}
headers <- c("fawn","antelope","precipitation","winter")
colnames(df) <- headers
str(df)
```

4. Create	bivariate	plots	of	number	of	baby	fawns versus	adult	antelope	population,	the	precipitation	that	year,	and	the	severity	of	the	winter.	Your	code	should	produce	three	separate	plots.	Make	sure	the	Y-axis	and	X-axis	are	labeled.	Keeping	in	mind	that	the	number	of	fawns is	the	outcome	(or	dependent)	variable,	which	axis	should	it	go	on	in	your	plots?

```{r}
ggplot(data = df) + geom_point(aes(x=antelope,y=fawn),size=3)
```

```{r}
ggplot(data = df) + geom_point(aes(x=precipitation,y=fawn),size=3)
```

```{r}
ggplot(data = df) + geom_point(aes(x=winter,y=fawn),size=3)
```

5. Next,	create	three	regression	models	of	increasing	complexity	using	lm().	In	the	first	model,	predict	the	number	of	fawns	from	the	severity	of	the	winter.	In	the	second	model,	predict	the	number	of	fawns from	two	variables	(one	should	be	the	severity	of	the	winter).	In	the	third	model	predict	the	number	of	fawns from	the	three	other	variables.	Which	model	works	best?	

Which	of	the	predictors	are statistically	significant	in	each	model? If	you	wanted	to	create	the	most parsimonious	model	(i.e. the	one	that	did	the	best	job	with	the	fewest	predictors),	what	would	it	contain?

Model 1
```{r}
model1 <- lm(data = df, formula = fawn ~ winter)
summary(model1)
```
"winter" is statistically significant in this model. 


Model 2
```{r}
model2 <- lm(data = df, formula = fawn ~ winter + antelope)
summary(model2)
```
"antelope" is statistically significant in this model.

Model 3
```{r}
model3 <- lm(data = df, formula = fawn ~ .)
summary(model3)
```
All three variables ("antelope", "precipitation", and "winter") are statistically signfiicant in this model. 

Find the most important variable.
```{r}
#code below found from http://rstatistics.net/variable-importance-of-predictors-that-contribute-most-significantly-to-a-response-variable-in-r/
relImportance <- calc.relimp(model3, type = "lmg", rela = TRUE)
sort(relImportance$lmg, decreasing = TRUE)
```
Based on the above analysis, if I were to create the most parsimonious model I would just use the antelope variable since it has the highest relative importance. 
