---
title: "IST687_Lab3"
output:
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---

####Task 1: Write, test, and submit the necessary code in R to accomplish the following

1. Generate a normal distribution, or 1000 samples, with a mean of 80
```{r}
norm <- rnorm(1000,80)
```

2. Write a	function	that	takes	three	variables	- a	vector,	a	min	and	a	max,	and	returns	the	number	of	elements	in	the	vector	that	are	between	the	min	and	max (including the	min	and	max)
```{r}
numb <- function(vec, mini, maxi) {
  count <- 0
  for (x in vec) {
    if (mini <= vec[x] & vec[x] <= maxi) {
      count <- count + 1
    }
  }
  return(count)
}
```

3. Use	the	function	to	see	how	many	of	your	normal	distribution samples are	within	the	range	of	79	to	81
```{r}
numb(norm,79,81)
```

4. Repeat	3	times	(creating	a	normal	distribution	and	then	calling	your	function),	to	see	if	the	results	vary
```{r}
norm1 <- rnorm(1000,80)
norm2 <- rnorm(1000,80)
norm3 <- rnorm(1000,80)

numb(norm1,79,81)
numb(norm2,79,81)
numb(norm3,79,81)
```


####Task	2:	Write,	test,	and	submit	the	necessary	code	in	R	to	accomplish	the	following: 

```{r}
#Need to install the VGAM library to use for Pareto distributions
install.packages("VGAM")
library(VGAM)
```

1. Generate	51	random	numbers	in	a	Pareto	distribution	and	assign	them	to	a	variable	called	"FSApops."
```{r}
FSApops <- rpareto(51, 40000, 15)
FSApops
```

2. Specify	a	"location"	and	a	"shape"	for	your	Pareto	distribution	that	makes	it	as	similar	as	possible	to	the	actual	distribution	of	state	populations.
```{r}
FSApops1 <- rpareto(51, 40000, 30)
FSApops2 <- rpareto(51, 40000, 5)
FSApops3 <- rpareto(51, 40000, 50)
```

3. 	Create	a	histogram	that	shows	the	distribution	of	values	in	FSApops
```{r}
hist(FSApops) #FSApops appears to be the closest to the US state distribution with a couple outliers
hist(FSApops1)
hist(FSApops2)
hist(FSApops3)
```

4. 	Use	a	command	to	report	the	actual	mean	and	the	actual	standard	deviation	of	the	51	values	stored	in	FSApops.
```{r}
mean(FSApops)
sd(FSApops)
```

5. 	Use	a	command	to	report	the	population	of	your	largest	fictional	state	(i.e.,	your	California)	and	your	smallest	fictional	state	(i.e.,	your	Wyoming).
```{r}
max(FSApops)
min(FSApops)

```

